---
title: "Musicology"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
runtime: shiny
---

```{r global, include=FALSE}
# load data in 'global' chunk so it can be shared by all users of the dashboard
library(tidyverse)
library(spotifyr) 
library(ggplot2)
library(hrbrthemes)
library(compmus)


playlist <- get_playlist_audio_features('', '2EUygq17yMPzSbM6XTm4rs') 
playlist$release <- as.Date(playlist$track.album.release_date, format = "%Y-%m-%d")


```

```{r}
# Reactive that returns the whole dataset if there is no brush
selectedData <- reactive({
  data <- brushedPoints(playlist, input$plot_brush)
  if (nrow(data) == 0)
    data <- playlist
  data
})
```

# Introduction

## Column {data-width="650"}

### Raw data {data-width="600"}

```{r}
ggplot(playlist, aes(x=release, y=tempo)) +
 geom_point()
```

### Data processing

above you see a quick overview of the release date of the tracks in my playlist together with the tempo. One thing to note here is that the tempo of a lot of tracks is wrong. My guess is that spotify thinks the tempo is half of what it should actually be, since the tempo of tracks in my playlist should actually range between 140-200+.

This together with some of the outliers made before 2015 makes the graph look messy. so i decided to remove every entry from before 2015, and multiply every tempo that is lower than 135 by 2, which made the data a lot cleaner as you will see on the Interactive plot page

## Column {data-width="350"}

### Info {data-width="400"}

Fill in when finished, give info about where to see what

# Interactive plot

```{r include=FALSE}
specified_date <- as.Date("2015-01-01")
playlist <- playlist[playlist$release >= specified_date,]

playlist <- playlist[complete.cases(playlist$tempo), ]
playlist$tempo[playlist$tempo < 135] <- playlist$tempo[playlist$tempo < 135] * 2

playlist <- playlist[, c('track.name', 'tempo', 'energy', 'danceability', 'release')]
```

## Column {data-width="650"}

### Tempo vs release date (try selecting some tracks) {data-width="600"}

```{r}

renderPlot({
  ggplot(playlist, aes(x=release, y=tempo)) + geom_point()
}, outputArgs = list(brush = brushOpts(id = "plot_brush")))
```

### tempo vs energy vs danceability

```{r}
renderPlot({
  ggplot(selectedData(), aes(x=energy, y=tempo, color=danceability)) + 
    geom_point() +
    geom_smooth(method="lm", formula= (y ~ exp(2*x)), se=FALSE, color='red') +
    scale_color_gradient(low = "blue", high = "yellow")
})
```

## Column {data-width="350"}

### Selected tracks {data-width="400"}

```{r}
renderTable({
  selectedData()
})
```

# Chroma Features

```{R}
wood <-
  get_tidy_audio_analysis("38KNGUbQYgeLrOCgUK8vyx") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)
```




## Column {data-width="650"}

### Chromagram {data-width="600"}
```{R}
wood |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma() |> 
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = pitch_class,
      fill = value
    )
  ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude") +
  theme(aspect.ratio=9/16, legend.position="bottom") +
  scale_fill_viridis_c()
```

### Info
For the chromagram i decided to pick a track with a very euphoric melody. you can probably see some nice patterns in this graph but i know nothing about music theory so it doesnt really say anything to me
